\section{Threats to Validity}

Regarding construct validity,  we are actually relying on the subjective perception of
practitioners when we stated that we performed our study considering successful cases
of DevOps adoptions. However, currently, there is no objective way to measure whether or not a
DevOps adoption was successful.
Although Grounded Theory offers rigorous procedures for data analysis, some degree of research bias
might have occurred in our qualitative research. Certainly other researchers might form a different
interpretation and theory after analyzing the same data,  but we believe that the main perceptions
would be preserved. This is a typical threat related to GT
studies, which do not claim to generate definitive findings. The resulting theory, for instance, might
be different in other contexts \cite{hoda2012developing}.

For this reason, we do not claim
that our theory is absolute or final. We welcome extensions to the theory based
on unseen aspects or finer details of the present categories or potential discovery
of new dimensions from future studies.
Future work can also focus on investigating contexts
where DevOps adoption did not succeed, aiming to validate if our model could be
relevant in this scenario too. Finally, regarding external validity, although we
considered in our study the point of view of practitioners with different
backgrounds, working in companies from different domains, and distributed across
five countries, we do not claim that our results are valid for
other scenarios---although we almost achieved saturation
after the 12th interview. Accordingly, our degree of heterogeneity complement
previous studies that mostly focus in a single company (as we discuss next).
