\section{Related Work} \label{sec:related_work}

The research literature is particularly rich when it comes to DevOps-related
works (e.g.,~\cite{devops_a_definition_xp_15,extending_dimensions_icsea_16,qualitative_devops_journalsw_17}).
In a literature review, Erich et al.~\cite{cooperation_dev_ops_esem_2014} presents 8
main concepts related to DevOps: culture, automation, measurement, sharing,
services, quality assurance, structures and standards. The authors pointed out
that the first four concepts are
related to the CAMS framework, proposed by Willis~\cite{what_devops_means_2010}.
The paper concludes that there is a great opportunity for empirical researchers
to study organizations experimenting with DevOps.
Other studies (e.g.,~\cite{devops_a_definition_xp_15,dimensions_of_devops_xp_15,extending_dimensions_icsea_16,characterizing_devops_sbes_2016,qualitative_devops_journalsw_17})
mixed literature reviews with empirical data to investigate DevOps.
Although our research and recent literate are interested in understanding DevOps,
there are subtle differences in both (1) the methodological aspects and (2) the focus
of each work.

First of all, none of the aforementioned works focused on explaining the process of DevOps adoption,
in particular, using data collected in the industry. This is unfortunate, since the
practitioners' perception present an unique point of view that researchers
alone could hardly grasp. Moreover, although the literature has a number of
useful elements, there is a need to complement such elements with a perspective on how DevOps has
been adopted, containing guidance about how to connect all these isolated parts
and then enabling new candidates to adopt DevOps in a more consistent way.
For instance, the work of Erich et al.~\cite{qualitative_devops_journalsw_17}
focus on investigating the ways in which organizations implement DevOps.
However, this work relies only in literature review and does not formulate
new hypothesis about DevOps adoption. Second,
in terms of results, our main distinct contribution is to improve the guidance
to new practitioners in DevOps adoption.
Next, we present the overlappings of our
results with the existing literature, presenting also the main differences that
make the contributions of our work clearer.

The work of Smeds et al.~\cite{devops_a_definition_xp_15} uses a literature
review to produce one explanation about DevOps through a set of enablers and capabilities. Additionally, their results
present a set of impediments of DevOps adoption based on an interview with 13
subjects of a same company, and whose DevOps adoption process was at
an initial stage. The main similarities with our study are: (1) grouping
elements as DevOps enablers; and (2) the presence of several similar concepts:
(a) testing, deploying, monitoring, recovering and infrastructure automation;
(b) continuous integration, testing and deployment; (c) service failure recovery
without delay; and (d) constant, effortless communication. The main
differences are: (1) their work does not group concepts into categories,
for example: most of their enablers were grouped together by us within the \cat{automation} category; (2) presents cultural enablers as
common contributor to DevOps, not as the most important concern; and (3) the empirical
part of the study focus on building a list of possible impediments to DevOps
adoption, not on providing guidance to new adopters.


In the study of Lwakatare et al.~\cite{dimensions_of_devops_xp_15}, the
authors aimed at characterizing the and formalize what DevOps is about. Through
a sequence of interviews, the authors observed the need of four dimensions to compose
DevOps, including collaboration, automation, measurement, and monitoring.
In a follow up study,
Lwakatare et al.~\cite{extending_dimensions_icsea_16} proposed a conceptual
framework to explain ``DevOps as a phenomenon''. The framework is organized around
five dimensions (collaboration, automation, culture, monitoring and measurement)
and these dimensions are presented with related practices.
These two works have good similarities with our study. For instance,
all aforementioned dimensions are also presented here. The
main differences are: (1) collaboration and culture are presented by us
as a single abstraction; (2) Concepts related to monitoring and measurement are
grouped by us in a single category: \cat{continuous measurement}; and (3) it does
not indicate a major dimension (aka, the core category).
Moreover, our work greatly expand the notion of DevOps, proposing a theory
for adoption, and indeed applying this theory in a real setting.

Fran\c{c}a et al.~\cite{characterizing_devops_sbes_2016} present a DevOps
explanation produced by means of a multivocal literature review. The data was collected
from multiple sources, including gray literature, and analyzed by using procedures
from GT. The results contain a set of DevOps principles, where
there is most of the overlapping with our study. In addition, the paper
presents a definition to DevOps, issues motivating its adoption, required skills,
potential benefits and challenges of adopting DevOps. The main similarities
are: (1) Automation, sharing, measurement and quality assurance are presented as
DevOps categories; and (2) Their social aspects category is similar to our
\cc category. The main differences are: (1) it presents DevOps as a
set of principles, different from enablers and outcomes in our study; and (2) the Leanness
category is not present in our study and the \cat{resilience} category is not present
in theirs; and (3) it does not indicate a core category.

The study conducted by Erich et al.~\cite{qualitative_devops_journalsw_17},
similarly to the others cited above, combined literature review with some
interviews with practitioners. In the literature review part, the papers were
labeled and the similar labels are grouped. The 7 top labels are then presented
as elements of DevOps usage in literature: culture of collaboration, automation,
measurement, sharing, services, quality assurance and governance. After the literature
review, six interviews were conducted in order to obtain evidence of DevOps
adoption in practice. The interviews were analyzed individually and a comparison
between them was made, focusing on problems that organizations try to solve by
implementing DevOps, problems encountered when implementing DevOps and practices
that are considered part of DevOps. The main similarity with our study
is that 5 of their 7 groups are also present in our study (culture of collaboration,
automation, measurement, sharing and quality assurance). The main
differences are: (1) it does not consolidate the practitioners' perspective, but
only compare it with literature review results; and (3) it does not indicate a major group.

Finally, the work of Vergori and colleages~\cite{Vergori:2017:ICPE}, the
authors proposed a set of metrics related to DevOps performance, such as
expected task completion rate, expected finishing time, and the proportion of
time doing dev or ops activities. Although some of the metrics proposed are
reasonable straightforward to measure (e.g., the task completion rate), some
other are not so easy. In this work, the authors used the Phoenix project~\cite{Kim:2013:DevOps}, which
is a case of industrial DevOps adoption. In terms of similarities to our work,
both works focus on improving DevOps experience. While we focus on DevOps
adoption, their work focus on DevOps performance.

In comparison with our previous paper~\cite{Luz:2018:ESEM}, here we advance
further in exploring real scenarios of DevOps usage. The TCU scenario was
described in details through the results of the focus group. Besides that, we
present a more detailed explanation about the application of our model,
highlighting each step and presenting some numbers.
