\begin{abstract}

  \emph{ Research Context.} DevOps is a set of practices that has gained increasing
  attention in the industry and that aims to reduce the
  barriers between development and operations
  teams. Due to its increasing interest and imprecise
  definitions, existing research works have tried to
  characterize DevOps, mainly using a set of concepts and related practices.

  \emph{Research problem and goal.} However, little is
  known about the \emph{practitioners understanding}
  about successful paths for DevOps adoption, which might
  hinder institutions to start the use DevOps practices. Our goal
  here is to present a new theory about DevOps, highlighting the
  main related concepts that contribute to its adoption in industry. 

  \emph{Research Method.} Our work builds upon the Classical Grounded
  Theory approach. We considered the opinion of practitioners
  that contributed to the adoption of DevOps in 15 companies from different
  domains and across five countries. We empirically evaluate our model through
  a case study, whose goal is to increase the maturity of
  DevOps adoption at the  Brazilian Federal Court of Accounts,
  a Brazilian Federal Government institution. 

  \emph{Results.} This paper 
  presents a model to improve both the understanding and guidance
  of DevOps adoption. The model increments the existing view of
  DevOps by explaining the role and motivation of each
  element (and their relationships) in the DevOps adoption process.
  We organize this model in terms of \emph{DevOps enabler categories} and
  \emph{DevOps outcome categories}, and explain that \emph{collaboration}
  is the core DevOps concern, contrasting with an existing wisdom that
  \emph{automating building and deployment processes} is enough to
  achieve DevOps. 
  
  \emph{Implications.} Altogether, our results
  contribute to (a) generating an adequate understanding of DevOps,
  from the perspective of
  practitioners; and (b) assisting other institutions in the
  migration path towards DevOps adoption.
  
\end{abstract}

% no keywords
