\documentclass{letter}
\usepackage{hyperref}
\signature{Welder Luz}
\address{St. de Administra\c c\~{a}o Federal Sul \\ Asa Sul, Brasília - DF \\ 70042-900, Brazil}

\begin{document}

\begin{letter}{}
\opening{Dear JSS Editors}

We are submitting a new version of our manuscript
\emph{DevOps in the Real World: A Theory, a Model and a Case Study
  Journal of Systems and Software}, under review number JSS-D-18-01011.
In this new version we present the following major changes

\begin{itemize}
\item We introduced a GQM definition for our study (in the
  beginning of Section 2).
  
\item We introduced several
transcriptions of the interviews (Section 3 and Section 4).

\item We reviewed the description of our theory, and we
  describe it using a well-known framework for building
  theories on software engineering (Section 4.3).

\item We highlight that the current status of DevOps adoption
  at TCU enabled a more agile approach for system deployments.
  This benefit is what we can better quantify right now. We also
  briefly discuss that we were able to almost eliminate
  the downtime related to deployment activities, for a set
  of applications. We address this issue in Section 2 and
  Section 5. 

\item We make more clear the differences between this work
  and our previous publication at ESEM 2018. We address
  this issue in Section 1. In summary, in this version we
  review our theory, we describe our theory using a
  specific framework, and we report the status of DevOps
  adoption at TCU, using a study based on the focus group
  method.

\end{itemize}

We also addressed all minor comments as well, and we
would like to thank the anonymous reviewers to their
comments that helped us to improve the paper. 

\closing{Yours Faithfully,}

\end{letter}
\end{document}
